---
title: "DataSci 306 Final Project"
author: "DataSci 306 Instructional Team"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Investigating the Internet Movie Database (IMDB)

The [Internet Movie Database (IMDb)]() contains information on millions of movies and television programs. They offer several [non-commercial use datasets](https://developer.imdb.com/non-commercial-datasets/) (documentation link). For this project we will analyze a **sample** of 100,000 titles from the IMDBb. 


## Part I: Preprocessing

* [Edit your `.gitignore` file](https://docs.github.com/en/get-started/getting-started-with-git/ignoring-files) to ignore all files with the `.rda` extension. (Add and commit)
* Create a new file in the `data/` directory called "Preprocessing.Rmd". The remaining instructions in this section are to be completed in that file.
* Write a function that will load a table from the IMDb files in the `data/` directory.
  * The function should take the file name (without the ".csv.gz" portion) as an argument
  * The function should load the appropriate `.csv.gz` file.
  * Make sure that all "\\N" values (which IMDB uses to indicate missing values) are turned into proper NA values in R
  * The function should return the table.
* For each of the `.csv.gz` files, use your function to load the table, then save it into a variable (e.g. `name_basics <- preprocess("name_basics")`) and use the `write_rds` function (e.g., `write_rds(name_basics, "name_basics.rda")`.
* Run the function on all of the `*_sample.csv.gz` files to created processed `.rda` files.
* In your other files, you can load these using the `TABLE <- read_rds("data/FILENAME.rda")` function.

```{r}
# Example code for loading and saving a table
preprocess <- function(filename) {
  # Load the table
  table <- read_csv(paste0("data/", filename, ".csv.gz"), na =
                     c("\\N"))
  return(table)
}
name_basics <- preprocess("name_basics_sample")
write_rds(name_basics, "data/name_basics.rda")
titles_basics <- preprocess("title_basics_sample")
write_rds(titles_basics, "data/title_basics.rda")
ratings <- preprocess("title_ratings_sample")
write_rds(ratings, "data/ratings.rda")
principals <- preprocess("title_principals_sample")
write_rds(principals, "data/principals.rda")
```
## Part II: EDA of individual tables

* For each of the 4 tables, perform basic exploratory data analysis. Report the following information:
  * For each quantitative column, provide some summary statistics
  
```{r}
summary(titles_basics$runtimeMinutes)
```
  * For any character columns, decided if they are actually representing factors/categorical data with a moderate number of columns. If so report the distributions for these variables.
  
```{r}
  # Example code for categorical variables
  table(titles_basics$genres)
```
  * Provide a plot for each table. Across all of the plots, try to show off the most possible different ggplot features (`geoms_` functions, `stat_` functions, coordinate systems, facets, use of several variables, annotations)
  
```{r}
ggplot(titles_basics, aes(x = runtimeMinutes)) + geom_histogram()
```
